Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-19 main.Rmd
minor corrections

@@

```{r tab-mayan}
table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
                      K = c(20, 18, 20, 20, 1)) 
kable(table_mayan, 
      caption = "Hierarchy table for Mayan calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All cyclic granularities are circular with $K$ as the period length.")
```

@@

**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from Gregorian calendar. Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities. Only one single-order-up granularity day-of-month is quasi-circular.

|  G	        |C                 |K
|------------	|------------------|----------------|
|  minute     |minute-of-hour    |  60            |
|  hour    	  |hour-of-day       |  24  	        |
|  day      	|day-of-month      | k(day, month)  |
|  month    	|month-of-year     |  12            |
|  year   	  | 1                |	1             |

@@

```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure", fig.pos="htb"}
include_graphics("Figs/data-structure-v1.png")
```

@@

```{r search_gran_limit2, echo=FALSE, eval = FALSE}
smart_meter10 %>% search_gran(
  highest_unit = "month",
  filter_out = c("hhour", "fortnight")
)%>% knitr::kable(format = "markdown")
```

@@

```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot and (c) box plot of energy consumption for customer id: 10017936. Plot (a) helps to compare different hours of the day within weekdays and weekends. Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable due to different seasons. Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour."}

@@

First, it would be interesting to see if the distribution of total runs vary depending on if a team bats first or second? Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015. It would be interesting to take their example in order to dive deeper into this question. <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings. This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable. To answer Q1, winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.

@@

Another interesting question could be if runs per over set to reduce in subsequent over if fielding is good in the previous over? For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over. <!--If a batsman is bowled out, it does not necessarily signify good fielding. So we only include number of catches and run out in an over as a measure of good fielding. 
Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs. --> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable. Figure \ref{fig:field} shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.

>>>>

<!-- # ```{r tab-mayan} -->
<!-- # table_mayan <- tibble(`Circular granularity` = c("minute-of-hour","minute-of-day", -->
<!-- #                                                  "hour-of-day","hour-of-week", "day-of-week"), -->
<!-- #                       Expression = c($L_1(z \mod 60)$, L_2(z \mod 60*24), L_3(\lfloor z/60\rfloor\mod 24)$,  L_4(\lfloor z/60\rfloor\mod 24*7)$, L_5(\lfloor z/24*60\rfloor \mod 7)$)) -->
<!-- #                                                   -->
<!-- #                                                   -->
<!-- #                       C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1), -->
<!-- #                       K = c(20, 18, 20, 20, 1))  -->
<!-- # kable(table_mayan,  -->
<!-- #       format = "latex", -->
<!-- #       caption = "Hierarchy table for Mayan calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All cyclic granularities are circular with $K$ as the period length.") -->
<!-- # ``` -->

@@

```{r tab-mayan}
table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
                      K = c(20, 18, 20, 20, 1)) 
kable(table_mayan, 
      format = "latex",
      caption = "Hierarchy table for Mayan calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All cyclic granularities are circular with $K$ as the period length.")
```

@@

**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}. Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities. Only one single-order-up granularity day-of-month is quasi-circular.

```{r tab-gregorian}
table_greg <- tibble(G = c("minute", "hour", "day", "month", "year"),
                      C = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
                     K = c(20, 18,"k(day, month)", 20, 1))

kable(table_greg, 
      caption = "Illustrative hierarchy table for Gregorian calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All these cyclic granularities are circular with $K$ as the period length except for day-of-month which is quasi-circular.")
```

@@

```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
include_graphics("Figs/data-structure-v1.png")
```

@@

```{r search_gran_limit2, echo=FALSE, eval = FALSE}
smart_meter10 %>% search_gran(
  highest_unit = "month",
  filter_out = c("hhour", "fortnight")
)%>% knitr::kable()
```

@@

```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot and (c) box plot of energy consumption for customer id: 10017936. Plot (a) helps to compare different hours of the day within weekdays and weekends. Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable due to different seasons. Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour.", out.width="90%"}

@@

First, it would be interesting to see if the distribution of total runs vary depending on if a team bats first or second? Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015. It would be interesting to take their example in order to dive deeper into this question. <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings. This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable. To answer Q1, winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.

@@

Another interesting question could be if runs per over set to reduce in subsequent over if fielding is good in the previous over? For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over. <!--If a batsman is bowled out, it does not necessarily signify good fielding. So we only include number of catches and run out in an over as a measure of good fielding. 
Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs. --> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable. Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
